### 2021-11-23 paper.Rmd
some unordered stuff

Anticlustering methods can be distinguished by the input data (pairwise dissimilarity vs. attributes values) and the objective function that is used to optimize group between-group similarity or within-group heterogeneity. Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance. The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, i.e., the total sum of dissimilarities between any elements within the same group. Maximizing the diversity simultaneously optimizes within-group heterogeneity and between-group similarity. Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity. Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019]. Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.

Oftentimes, researchers work with attribute values [or *features*; @dry2009] instead of dissimilarity ratings. This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].  For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982]. In other cases, attribute values are binary and may represent the presence or absence of a feature [@tversky1977]. In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019]. Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups. A different approach to anticlustering directly works with the attribute data by maximizing the k-means criterion. K-means is probably the best-known clustering method. Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983]. In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006]. Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np]. For the anticlustering application, the k-means criterion is maximized instead, thereby achieving similarity between groups. Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters. Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored. When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973]. @papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.

@@

```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all partitions (for $N = 14$ and $K = 2$). The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}

@@

To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion. All methods used a standardization of the data sets as a preprocessing step. In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.

@@

In this paper, I presented the k-plus criterion for anticlustering. Anticlustering methods are used to divide sets of elements into subgroups while striving for high between-group similarity or high within-group heterogeneity. The k-plus criterion represents between-group similarity as the discrepancy between the mean and variance of attribute values. These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided. The k-plus criterion extends the classical k-means criterion---which represents how similar groups are with regard to their means---by incorporating a term that represents how groups differ with regard to their variances. The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable. Thus, the original k-means criterion remains basically unchanged. A simulation study and examples on norming data indicated that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.

>>>>

Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity or within-group heterogeneity. Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance. The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, which is the total sum of dissimilarities between any elements within the same group. Maximizing the diversity optimizes within-group heterogeneity, which simultaneously leads to similar groups---if all groups have the same size. Anticlustering objective functions usually represent the reversal of a clustering objective, only the direction of the optimization is changed. Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity. Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019]. Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.

Oftentimes, researchers work with attribute values [or *features*; @dry2009] instead of dissimilarity ratings. This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].  For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982]. In other cases, attribute values are binary and may represent the presence or absence of a feature [@tversky1977]. In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019]. Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups. A different approach to anticlustering directly works with the attribute data by maximizing the k-means criterion. K-means is probably the best-known clustering method. Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983]. In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006]. Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np]. For the anticlustering application, the k-means criterion is maximized instead, thereby achieving similarity between groups. Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters. Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored. When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973]. @papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on means.

@@

```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups. The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}

@@

To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion. All methods used a $z$-standardization of the data sets as a preprocessing step. In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.

@@

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on constructing groups that are highly similar to each other, operationalized by similar means as well as variances. These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data. For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable. A simulation study and examples on real norming data indicate that minimizing differences with regard to means and variances between groups is well possible using the k-plus criterion.
